---
title: "causal_01_causenet_prep.Rmd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Parsing CauseNet Precision JSON

The CauseNet precision dataset was downloaded from [this link](https://groups.uni-paderborn.de/wdqa/causenet/causality-graphs/causenet-precision.jsonl.bz2) on 7/14/2021 and extracted with 7zip. The extracted `causenet-precision.jsonl` file was placed in the `data_in` subdirectory.

```{r}
library(jsonlite)
library(readr)
library(tibble)

line_to_df <- function(X) {
    json <- jsonlite::fromJSON(X)
    tibble(
        cause = json$causal_relation$cause$concept,
        effect = json$causal_relation$effect$concept,
        sentence = json$sources$payload$sentence
    )
}
```

The json is parsed to produce a dataframe row for each sentence. 
An intermediate csv file is streamed to the `scratch` subdirectory.

```{r}
cn_file <- "data_in/causenet-precision.jsonl"
N <- 1
N_max <- 200000
f <- file(description = cn_file, open = "r")
outfile <- "scratch/causenet_parsed.csv"
unlink(outfile)
read <- TRUE
while (read) {
    if (N %% 50 == 0) print(sprintf("Prop done: %.3f", N/N_max))
    N <- N + 1
    X <- readLines(f, n = 1)
    if (N  >= N_max) break
    if (is.null(X)) break
    Y <- line_to_df(X)
    write_csv(Y, outfile, append = TRUE)
}
```

Encountered an error at N = 197808, at 98.9% percent done

```{r}
cn_file <- "data_in/causenet-precision.jsonl"
N <- 197809
N_max <- 200000
f <- file(description = cn_file, open = "r")
outfile <- "scratch/causenet_parsed.csv"
read <- TRUE
while (read) {
    if (N %% 50 == 0) print(sprintf("Prop done: %.3f", N/N_max))
    N <- N + 1
    X <- readLines(f, n = 1)
    if (N  >= N_max) break
    if (is.null(X)) break
    Y <- line_to_df(X)
    write_csv(Y, outfile, append = TRUE)
}
```

Remaining rows were appended without issue

## Subsample Parsed CauseNet Data

Load the outputted csv back in and do a seeded subsample of 500,000 rows.
After deduplication, there is ~946k unique entries, so this is roughly half.
The subsampled csv is exported to the `scratch` subdirectory.

```{r}
library(dplyr)

prec <- read_csv("scratch/causenet_parsed.csv",
                 col_names = c("cause", "effect", "evidence"))

prec_2 <- distinct(prec)
```

```{r}
set.seed(4313469) # did sample(1:10000000, 1) to get a random seed
prec_2 %>%
    sample_n(500000) ->
    prec_sub

write_csv(prec_sub, "scratch/causenet_precision_500k.csv")
```


